R Notebook

### 2018-09-18 study2/lab_pres_nb_2018-6-11.Rmd
added files

```{r load in existing objects, message=FALSE, warning=FALSE}
# net.struct <- readRDS("net-struct-cogsci2018.rds")

source("../../study1/vacc_import_data.R")
source("../../Scripts/gmod_tools.R")
# source("../../Scripts/virtualEvidence.R")

library(bnlearn)
library(rjags)
library(patchwork)
library(BiDAG)
library(HydeNet)
library(bnlearn)

source("cog-model-main.R")
```

```{r set up training split}
# rescale beta

rescale_beta <- function(x, lower, upper) {
  # rescales onto the open interval (0,1)
  # rescales over theoretical bounds of measurement, specified by "upper" and "lower"

  N <- length(x)
  res <- (x - lower) / (upper - lower)
  res <- (res * (N - 1) + .5) / N

  return(as.vector(res))
}

d_bn_scaled <- d_bn %>%
 mutate_all(function(x){rescale_beta(x,-3,3)})

## set the seed to make your partition reproductible
set.seed(123)
trainInd <- sample(seq_len(nrow(d_bn_scaled)), size = floor(nrow(d_bn_scaled)*.80))

train <- d_bn_scaled[trainInd, ]
test <- d_bn_scaled[-trainInd, ]
```

# loading in data model

```{r}
# first load in mcmc results
all_results <- readRDS("allresults.rds")
nodes <- colnames(train)
# then grab map dag for theory-based constraints
merged_chains <- merge_chains(all_results[[5]])
map_dag <- extract_mapdag(merged_chains)

h <- HydeNetwork(as.formula(bnlearn_to_hyde_string(map_dag)))
plot(h)
```

## creating cognitive model

```{r}
# parse bn into list of parents and children

arcs_df <- bnlearn_to_df(map_dag)

df_to_list <- function(arcs_df, as.strings=FALSE) {
  # take arcs_df and make list of formulas
  arcs_df$to <- as.character(arcs_df$to)
  arcs_df$from <- as.character(arcs_df$from)

  nodes <- unique(unlist(arcs_df))
  output <- lapply(nodes, function(node) {
    edges <- arcs_df[which(arcs_df$to == node), "from"]
    if (length(edges) < 1) {
      edges <- NULL
    }
    list(child=node, parents=edges)

  })
  names(output) <- nodes
  return(output)
}

graph_list <- df_to_list(arcs_df)

bagOfModels <- lapply(graph_list, function(x){fit_node_cpt(x, train)})

# bagNet <- HydeNetwork(bagOfModels)
# writeNetworkModel(bagNet, pretty=TRUE)

```{r}
# add evidence node with cpt defined 
evid_ratio <- exp(0.4566215)

evid_cpt <- make_custom_cpt(c(.5*evid_ratio,.5, (1-.5*evid_ratio), (1-.5)), c(2,2), list(diseaseSevere=c("Yes","No"), evid = c("Yes","No") ), NULL)
# evid_cpt <- make_custom_cpt(c(.83,.5, .17, .5), c(2,2), list(diseaseSevere=c("yes","No"), evid = c("yes","no") ), NULL)
bagOfModels$evid <- evid_cpt

bagNet <- HydeNetwork(bagOfModels)
plot(bagNet)
writeNetworkModel(bagNet, pretty=TRUE)
## In the future (soon), make a function to create this node cpt

__Translating hydenet model to bnlearn__

below is some code to get things into the right format. Later I'll break this out into a function I guess.

```{r}
hyde_to_bn_cpt <- function(hyde_cpt) {
  ndims <- length(dim(hyde_cpt))

  if ("xtabs" %in% class(hyde_cpt)) {
    output <- hyde_cpt/1e5
  } else {

    output <- aperm(hyde_cpt, ndims:1)
  }
 return (output) 
}

# hyde_to_bn_cpt(bagOfModels$evid)

bnlearn_models <- lapply(bagOfModels, hyde_to_bn_cpt)

model_string <- as.character(map_dag)
model_string <- paste0(model_string,"[evid|diseaseSevere]")
model_network <- model2network(model_string)

predModel <- custom.fit(model_network, dist = bnlearn_models)
```

__Generating predictions__

Now I can use this model to generate predictions.

```{r}
# cpquery(predModel, event = (eval(parse(text="diseaseSevere"))=="Yes"), evidence = TRUE)
# cpquery(predModel, event = (eval(parse(text="diseaseSevere == 'Yes'"))), evidence = TRUE)
# cpquery(predModel, event = (eval(parse(text="diseaseSevere"))=="Yes"), evidence = (evid == "Yes"))

set.seed(12345)

pred0 <- sapply(nodes, function(x){
  statement <- paste0("cpquery(predModel, event = (", x, " == 'Yes'), evidence=TRUE, n = 5e5)")
  eval(parse(text=statement))
})

pred1 <- sapply(nodes, function(x){
  statement <- paste0("cpquery(predModel, event = (", x, " == 'Yes'), evidence= (evid == 'Yes'), n = 5e5)")
  eval(parse(text=statement))
})

pred_changes <- pred1 - pred0

pred_changes <- as.data.frame(pred_changes)

pred_changes <- pred_changes %>%
  mutate(scale = nodes) %>%
  rename(Mean = pred_changes) %>%
  mutate(type = "predicted")

```{r}
source("../../study3/load-study3-data.R", chdir = TRUE)

obs_changes <- d_scored %>%
  mutate(mean = rescale_beta(mean, -3, 3)) %>%
  spread(phase,mean) %>%
  mutate(changeScore = post-pre) %>%
  filter(condition == "diseaseRisk") %>%
  group_by(scale) %>%
  summarise(
    Mean = mean(changeScore),
    ul = mean(changeScore) + 1.96*sd(changeScore)/sqrt(n()), # replace with bootstrapping
    ll = mean(changeScore) - 1.96*sd(changeScore)/sqrt(n())
    ) %>%
  mutate(type = "observed")

obs_pred <- bind_rows(pred_changes, obs_changes) %>%
  spread(type, Mean) %>%
  group_by(scale) %>%
  summarize_all(mean, na.rm=TRUE)

cor(obs_pred$predicted, obs_pred$observed)

obs_pred %>% # a bit of a hack
  ggplot(aes(x=predicted, y = observed, ymin = ll, ymax=ul)) +
  geom_abline(slope=1, intercept=0, linetype="dashed", alpha =.5) +
  geom_point() +
  geom_errorbar() +
  coord_cartesian(xlim=c(-.06,.08), ylim=c(-.06,.08)) +
  theme_bw() +
  theme(aspect.ratio = 1)

obs_pred %>%
  gather(type, Mean, observed, predicted) %>%
  ggplot(aes(y = reorder(scale, Mean), x = Mean, xmin = ll, xmax = ul, color=type)) +
  geom_point() +
  geom_errorbarh(height=0)
```

----------------------------------------------

HydeNet prediction failures below ...

```{r}
set.seed(456)
model0 <- compileJagsModel(bagNet)
model1 <- compileJagsModel(bagNet, data= data.frame(evid=2))

post0 <- HydeSim(model0,
                       variable.names = nodes,
                       n.iter = 1e5
                       )

post1 <- HydeSim(model1,
                       variable.names = nodes,
                       n.iter = 1e5
                       )

post0 <- post0 %>%
  mutate(nat = ifelse(nat=="Yes",2,1)) %>%
  mutate_all(~ . -1) %>%
  summarize_all(mean) %>%
  select(-chain_index, -obs_index)

post1 <- post1 %>%
  mutate(nat = ifelse(nat=="Yes",2,1)) %>%
  mutate_all(~ . -1) %>%
  summarize_all(mean) %>%
  select(-evid, -chain_index, -obs_index)

pred_changes <- (post1 - post0)  %>%
  gather(scale, Mean) %>%
  mutate(type = "predicted")
```

```{r}
source("../../study3/load-study3-data.R", chdir = TRUE)

obs_changes <- d_scored %>%
  mutate(mean = rescale_beta(mean, -3, 3)) %>%
  spread(phase,mean) %>%
  mutate(changeScore = post-pre) %>%
  filter(condition == "diseaseRisk") %>%
  group_by(scale) %>%
  summarise(
    Mean = mean(changeScore),
    ul = mean(changeScore) + 1.96*sd(changeScore)/sqrt(n()),
    ll = mean(changeScore) - 1.96*sd(changeScore)/sqrt(n())
    ) %>%
  mutate(type = "observed")

bind_rows(pred_changes, obs_changes) %>%
  spread(type, Mean) %>%
  group_by(scale) %>%
  summarize_all(mean, na.rm=TRUE) %>% # a bit of a hack
  ggplot(aes(x=predicted, y = observed, ymin = ll, ymax=ul)) +
  geom_abline(slope=1, intercept=0, linetype="dashed", alpha =.5) +
  geom_point() +
  geom_errorbar() +
  # coord_cartesian(xlim=c(-.1,.1), ylim=c(-.1,.1)) +
  theme_bw() +
  theme(aspect.ratio = 1)
```

Strong correlation, and pretty good numerically. But for some reason it's getting disease severity wrong! I really can't wrap my head around it, it's not predicting what I would expect based on the evidence ratio (when I do it manually, the predictions are more in line with observed). So something strange is going wrong. I should try making a really simple little test case.

```{r}
bind_rows(pred_changes, obs_changes) %>%
  spread(type, Mean) %>%
  group_by(scale) %>%
  summarize_all(mean, na.rm=TRUE) %>%
  mutate(abs_resid = abs(predicted - observed)) %>%
  arrange(-abs_resid)
```

So for some reason this is not working quite right with hydenet. I'm going to translate things over to bnlearn and try taht.
# JUNK BELOW
----------------------------------------

```{r}

make_predictions_hyde <- function(hyde_model, orig_data, nodes_to_predict, nodes_held_out=c(), iter=1e3) {
  # takes a model and dataframe and generates model predictions for variables listed in 
  # nodes_to_predict. By default, uses all remaining nodes in orig_data. Hold out nodes from
  # prediction with nodes_held_out. iter sets mcmc samples

  pred_df <- orig_data
  pred_df[,nodes_to_predict] <- NA

  for (i in 1:nrow(pred_df)) {
    row <- pred_df[i,]
    row[nodes_held_out] <- NA
    pred <- do_jags_inference(hyde_model, nodes_to_predict, data = row, iter = iter)
    point_preds <- rowMeans(t(pred[[1]][(iter/2+1):iter,nodes_to_predict]))
    pred_df[i,nodes_to_predict] <- point_preds
  }

  return(pred_df)
}

predict_change_scores_cog <- function(model, nodes_to_predict, iter){

  preds_mean1 <- make_predictions(model, 
                                data.frame(evid=1), 
                                nodes_to_predict, 
                                c(),
                                iter = iter)

  preds_mean0 <- make_predictions(model, 
                                  data.frame(), 
                                  nodes_to_predict, 
                                  c("evid"),
                                  iter = iter)

  preds_mean <- preds_mean1 - preds_mean0
}

predict_change_scores_cog(bagNet, nodes, 1e4)
```

>>>>

```{r load in existing objects, message=FALSE, warning=FALSE}
net.struct <- readRDS("net-struct-cogsci2018.rds")

source("../study1/vacc_import_data.R")
source("../Scripts/gmod_tools.R")
source("../Scripts/virtualEvidence.R")

library(bnlearn)
library(rjags)
library(patchwork)
```

```{r load and tidy study 3 data, include = F}
# load data
d_all <- read.csv("data/study3_data.csv")

# reformat to match previous analyses (i.e., 2 rows per participant)
d_demo <- d_all %>% 
  select(workerId, condition, gender, age, ethnicity, education, job, income,
         political_party, political_beliefs, eligible_pretest, 
         is_parent_posttest, children_num_posttest, children_oldest_posttest, 
         children_youngest_posttest, plan_parent_posttest,
         starts_with("flushot_"), starts_with("vax_"), starts_with("attention_"),
         starts_with("comments"), starts_with("duration"))
d_pre <- d_all %>% 
  select(workerId, ends_with("_pretest")) %>%
  select(-c(eligible_pretest, starts_with("flushot_"), starts_with("vax_"), 
            starts_with("attention_"), starts_with("comments"), 
            starts_with("duration"))) %>%
  rename_all(funs(gsub("_pretest", "", .))) %>%
  mutate(phase = "pre")
d_post <- d_all %>% 
  select(workerId, ends_with("_posttest")) %>%
  select(-c(is_parent_posttest, children_num_posttest, children_oldest_posttest,
            children_youngest_posttest, plan_parent_posttest,
            starts_with("flushot_"), starts_with("vax_"), starts_with("attention_"),
            starts_with("comments"), starts_with("duration"))) %>%
  rename_all(funs(gsub("_posttest", "", .))) %>%
  mutate(phase = "post")

d <- bind_rows(d_pre, d_post) %>% 
  gather(question, response, -c(workerId, phase)) %>%
  mutate(phase = factor(phase,
                        levels = c("pre", "post")),
         reverse_cat = ifelse(grepl("_[1-9]r$", question), TRUE, FALSE),
         # NOTE: "response" has already been reverse coded!
         question = factor(question),
         scale = factor(gsub("_.*$", "", question),
                        levels = c("vaccIntent", "vaccDanger", "vaccEff", 
                                   "vaccStrain", "vaccTox", 
                                   "diseaseSevere", "diseaseRare", 
                                   "infantImmLimCap", "infantImmWeak", 
                                   "medSkept", "hb", "nat", 
                                   "overpar", "parentExpert"))) %>%
  full_join(d_demo) %>%
  mutate(condition = factor(condition, levels = c("noInterv", "diseaseRisk"))) %>%
  filter(!is.na(response), !is.na(workerId), !is.na(condition)) %>%
  distinct()

# how many left?
d %>% distinct(workerId, condition) %>% count(condition)

# score all scales
d_scored <- d %>%
  select(workerId, condition, phase, scale, response,
         gender:duration_posttest) %>%
  group_by(workerId, condition, phase, scale) %>%
  mutate(response = as.numeric(response)) %>%
  summarise(Mean = mean(response, na.rm = TRUE)) %>%
  ungroup() %>%
  distinct()

stderr <- function(x) {
          sqrt(var(x[!is.na(x)]) / length(x[!is.na(x)]))
}

interv_df <- d_scored %>% 
              filter(condition=="diseaseRisk") %>%
              mutate(interv = ifelse(phase=="pre",0,1)) %>%
              group_by(workerId) %>%
              spread(scale, Mean)

obs_data <- d_scored %>% 
  filter(condition == "diseaseRisk") %>% 
  spread(phase, Mean) %>%
  mutate(score = post - pre) %>%
  group_by(scale) %>%
  summarize(
    observed = mean(score),
    std_error = stderr(score))
```
